Structural Breaks in Inflation Dynamics within the European Monetary Union

### 2011-04-02 pkg/paper/Paper/draft.Rnw
commented comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@79 b679b2f7-6539-45d0-a605-5ddcb976cde9

\Abstract{
(... text when everything else finished, present innovation of paper...)
}

@@

(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
(... economic story, big transition ...) \\

@@

(... big transition continued, our contribution...) \\

@@

(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\

@@

(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\

@@

(... Jesus would present data later, first describe what we are going to to do...) \\
HICP series\footnote{seasonally unadjusted} from Jan.1990--Mar.2010 are obtained for 21 countries the \citet{oecd}. Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta are not included since they are very small economies. \\
The countries in this sample can be divided into three different groups. First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).  Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU. \\

@@

% laut achim noch zuviel bla bla im text --> weniger als anektode, präziser
(...less anectodic, more precise...) \\
To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks. First, the model is fitted for the whole sample using the generalized logistic distribution. This then enables testing for the stability of the parameter estimates. If there is evidence of significant changes, one proceeds by computing the number and the timing of the break points. \\
This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible. Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame. \\
(...discuss correlation...) \\
The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation. 
%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.

@@

(... formulation problems ...) \\
(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
%Motivation Verteilung
The data suggested that the use of a normal distribution might cause some estimation problems, since most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail. To accommodate for this fact, the use of the generalized logistic distribution (GL) with its relative low degree of complexity, seemed appropriate. On the one hand, it is a simple distribution and on the other hand, it is flexible enough to capture essential properties of the data. For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:

@@

\subsection{Test}
(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\

@@

\section{Power Analysis}

(...Jesus would not do this ...) \\

Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher. If for example, just the skewness parameter changes, is a test build on the normality assumption able to capture this? (...more to come if we decide to include this. In principle, we could then argue in section \ref{s31} that a power analysis suggested that we might miss changes in the third moment or so, i.e. normal has lower power if series is GL--distributed)...)

@@

(... perhaps put the data problems into appendix? Presently I deal with this in $results3.rnw$ in the following fashion: I do assume that for Belgium, Luxembourg and Portugal (the countries where the exact time and date of inclusion of sales effects is definitely known) the seasonality in these months follows this logic: DECREASE in prices in January, almost offset by an increase in February and another DECREASE in July followed by an increase in August. So the series AFTER the inclusion of the sales periods should take into account that the seasonal effects of these month are now HIGHER than before. I do this by substracting the difference between the seasonal dummy in January (prior to inclusion) and the seasonal dummy in January (after the inclusion) from the original series. If this is the best way to deal with that, I do not know ...) \\

@@

(... write when rest is finished, talk about method and empirical findings...)

>>>>

\Abstract{
%(... text when everything else finished, present innovation of paper...)
}

@@

%(... Jesus: journals, EE (empirical economics): mehr EMU, weniger Methode, Journal of Applied Econometrics: genau umgekehrt...)
%(... economic story, big transition ...) \\

@@

%(... big transition continued, our contribution...) \\

@@

%(...introduction needs to include the value added of this paper, why is it better, how does it compare to others...) \\

@@

%(...effects of monetary unions, literature overview (more literature in discussion of results)...) \\

@@

%(... Jesus would present data later, first describe what we are going to to do...) \\
HICP series\footnote{seasonally unadjusted} from Jan.1990--Mar.2010 are obtained for 21 countries the \citet{oecd}. Countries included are Austria,	Belgium, Czech Republic,	Denmark, Estonia,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,	Slovakia, Slovenia,  Spain,	Sweden, and	United Kingdom. Latvia and Lithuania as well as Bulgaria and Romania are excluded due to data scarcity. Cypria and Malta are not included since they are very small economies. \\
The countries in this sample can be divided into three different groups. First the euro countries (Austria, Belgium, Estonia, France, Finland, Germany, Greece, Ireland, Italy, Luxembourg, the Netherlands, Portugal, Slovakia, Slovenia and Spain).  Second, the EU members not participating in the ERM~II: Czech Republic, Hungary, Poland, United Kingdom, and Sweden. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU. \\

@@

% laut achim noch zuviel bla bla im text --> weniger als anektode, präziser
%(...less anectodic, more precise...) \\
To monitor changes in the dynamics of the inflation rates, we follow the usual strategy employed when testing for structural breaks. First, the model is fitted for the whole sample using the generalized logistic distribution. This then enables testing for the stability of the parameter estimates. If there is evidence of significant changes, one proceeds by computing the number and the timing of the break points. \\
This framework allows to trace the evolution of mean, variance and skewness of the inflation series and thus interpreting deviations in any of these parameters as possibly due to a change in monetary policy, is made feasible. Any huge intervention or change is captured by fluctuations in the parameter stability of the series and can be quantified and related to a distinct time frame. \\
%(...discuss correlation...) \\
The correlation dynamics of the series is of less importance to this framework, which concentrates on changes in mean, variance and skewness and does not focus on autocorrelation. 
%To provide consistent estimates of the parameter vector $\hat{\phi}$, the covariance matrix $\hat{V}^{-1/2}$ is estimated using a heteroskedasticity and autocorrelation consistent covariance estimator.

@@

%(... formulation problems ...) \\
%(...Jesus: GL not standard, explain why it is worth using it, what makes it better than ARIMA?...) \\
%Motivation Verteilung
The data suggested that the use of a normal distribution might cause some estimation problems, since most of the time series exhibit asymmetric properties (with sometimes rather strong skewness) and -- compared to the normal distribution -- a fatter tail. To accommodate for this fact, the use of the generalized logistic distribution (GL) with its relative low degree of complexity, seemed appropriate. On the one hand, it is a simple distribution and on the other hand, it is flexible enough to capture essential properties of the data. For the inflation series $y_t = 100\cdot log(HICP_t/HICP_{t-1})$ we therefore assume a GL distribution as defined in \citet{johnson}:

@@

\subsection{Test}
%(.. Jesus would cite a paper by Andrews and Ploberger 1993 that provided the key for creating confidence intervals around the estimates...) \\

@@

%\section{Power Analysis}
%
%(...Jesus would not do this ...) \\
%
%Basis idea: show that with a GL instead of a Normal, the probability of detecting minimal effects in the series (which is GL--distributed) is higher. If for example, just the skewness parameter changes, is a test build on the normality assumption able to capture this? (...more to come if we decide to include this. In principle, we could then argue in section \ref{s31} that a power analysis suggested that we might miss changes in the third moment or so, i.e. normal has lower power if series is GL--distributed)...)

@@

%(... perhaps put the data problems into appendix? Presently I deal with this in $results3.rnw$ in the following fashion: I do assume that for Belgium, Luxembourg and Portugal (the countries where the exact time and date of inclusion of sales effects is definitely known) the seasonality in these months follows this logic: DECREASE in prices in January, almost offset by an increase in February and another DECREASE in July followed by an increase in August. So the series AFTER the inclusion of the sales periods should take into account that the seasonal effects of these month are now HIGHER than before. I do this by substracting the difference between the seasonal dummy in January (prior to inclusion) and the seasonal dummy in January (after the inclusion) from the original series. If this is the best way to deal with that, I do not know ...) \\

@@

%(... write when rest is finished, talk about method and empirical findings...)
